---
title: "Juvenile snowshoe hares in Bonanza Creek Experimental Forest"
author: "Lucas Boyd"
date: "11/4/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# attach packages
library(here)
library(janitor)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(effsize)
library(broom)
library(equatiomatic)
```

## 1. Introduction and Overview

\*\*\*\*Insert 4-5 sentences of introduction

## 2. Data and Analyses

\*\*\*\*3-4 sentences of describing the data

```{r}
hares <- read_csv(here("data", "bonanza_hares.csv")) %>% 
  clean_names() %>%# cleaning names
  filter(age == "j") %>% # filtering for juveniles
  mutate(date = mdy(date)) %>% # converting date format
  mutate(year = year(date)) %>%  # creating a new year column
  mutate(sex = case_when( # changing names of observations
    sex == "f" ~ "Female",
    sex == "m" ~ "Male",))
```

## 3. Preliminary Results

#### A. Annual observations of juvenile hares

```{r}
hares_by_year <- hares %>% 
  count(year) # creating a subset for hare totals by year to explore the data
hares_by_year_mean <- mean(hares_by_year$n) # storing summary statistics
hares_by_year_max <- max(hares_by_year$n)
hares_by_year_min <- min(hares_by_year$n)
```

```{r}
ggplot(data = hares_by_year, aes(x = year, y = n)) + 
  geom_col(fill = "cyan4", width = 0.7) + # creating a bar graph of trap counts
  labs(x = "Year", y = "Total Juveniles Trapped") +
  scale_x_discrete(limit = c(1999:2012)) + # manually putting ticks for x axis
  theme_grey()
```

**Fig. 1** shows total the total number of juvenile snowshoe hares
trapped and observed each year across all sites.

As can be seen in this graph, number of juvenile trappings were not
consistent from year to year. The greatest number trapped was
`r hares_by_year_max` in 1999, the first year of the study. In 2002 and
2009, there were
`r`hares_by_year_min`juvenile observations collected. The mean number of hares trapped by year was`r
hares_by_year_mean\`. While this appears to show a downward trend in
juvenile hare population, the trend could also be a result of less
offort put into data collection. Going forward, researchers should be be
consistent about how many traps are set out and how often they're
checked to ensure accurate representations of snowshoe hare populations.

### B. Weight comparison and descriptive statistics

```{r}
# group and summarize the data for use in the visual
hares_weight_by_site_summary <- hares %>% 
  group_by(sex, grid) %>% 
  summarize(
    mean_weight = mean(weight, na.rm = TRUE))
```

```{r}
# visualizing average hare weight by site and sex
ggplot(data = hares_weight_by_site_summary, 
       aes(x = sex, y = mean_weight, fill = sex)) +
  geom_col(width = 0.8) +
  facet_wrap(~ grid) + # faceting by site
  labs(x = "Sex", y = "Mean Weight (g)") +
theme(legend.position = "none") +
  theme_grey() +
  theme(legend.position = "None") # removing the legend
  
```

**Fig. 2** compares the mean weight (g) of male and female juvenile
snowshoe hares across the three collection sites. "NA" indicates that
the sex was unspecified.

Across all three sites, male juvenile snowshoe hares had a higher mean
weight (g) than females and juvenile hares with unspecified sex. Both
mean female and mean male weight was the highest at site bonbs. This
seems to show that juvenile male snowshoe hares tend to be larger than
females. There were no observations with unspecified sex at this
location.

### C. Juvenile hare weight difference analysis

**Table 1** Descriptive statistics (mean, standard deviation, sample
size) of weight (g) of male and female juvenile snowshoe hares.

```{r}
hares_sex_table <- hares %>% 
  filter(sex %in% c("Male", "Female")) %>%
  drop_na(weight) %>% # filtering out unspecified sexes and weights
  group_by(sex) %>% # grouping by sex
  summarize( # finding summary statistics 
    mean_weight = round(mean(weight, na.rm = TRUE), 0), 
    sd_weight = round(sd(weight, na.rm = TRUE), 0), 
    sample_size = n())

hares_sex_table %>% # creating a stylized table with the stats
  kable(col.names = c("Sex", 
                      "Mean Weight (g)", 
                      "Standard Deviation (g)", 
                      "Sample Size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

```{r}
# storing vectors of male and female weights to use in t test

weight_male <- hares %>% # storing male weights
  filter(sex == "Male") %>% 
  drop_na(weight) %>% 
  pull(weight)


weight_female <- hares %>% # storing female weights
  filter(sex == "Female") %>% 
  drop_na(weight) %>% 
  pull(weight)
  
```

```{r, include = FALSE}
# exploratory plots to check out the data
hist(weight_female) # normally distributed, slightly positively skewed
hist(weight_male) # semi-normal, arguably bimodal
qqnorm(weight_female) # relatively linear
qqnorm(weight_male) # troublesome, not very linear
```

```{r}
# running a t.test of the weights
weight_by_sex_t <- t.test(weight_male, weight_female)

# finding and storing the actual difference in means
sex_weight_mean_difference = mean(weight_by_sex_t$estimate[1]) - 
  mean(weight_by_sex_t$estimate[2])

# finding and storing standard deviations for results
male_sd <- sd(weight_male)
female_sd <- sd(weight_female)

# finding and storing the effect size
sex_weight_effect_size <- cohen.d(weight_male, weight_female, na.rm = TRUE)
```

According to the results, juvenile male snowshoe hares (mean weight =
`r round(mean(weight_male), 0)` +/- `r round(male_sd, 0)`g) weigh
`r round(sex_weight_mean_difference, 0)`g more
(`r round((sex_weight_mean_difference)/(mean(weight_male)+mean(weight_female)/2)*100, 2)`%)
than females (mean weight = `r round(mean(weight_female), 0)` +/- `r round(female_sd, 0)`g). This
difference was determined to be significant (Welch's two sample t-test:
t( `r round(weight_by_sex_t$parameter, 3)`) =
`r round(weight_by_sex_t$statistic, 3)`, p =
`r round(weight_by_sex_t$p.value, 3)`), and the effect size is
small-moderate (cohens d =
`r round(sex_weight_effect_size$estimate, 3)`).

### D. Relationship between juvenile weight and hind foot length

```{r}
# creating a visual of weight vs. hind foot length
ggplot(data = hares, aes(x = hindft, y = weight)) +
  geom_jitter(color = "cyan4") +
  labs(x = "Hind Foot Length (mm)", y = "Weight (g)") +
  theme_grey()
```

\*\*\*\*add figure caption \*\*\*\*sexy up the graph

```{r}
# linear regression

hindft_weight_lm <- lm(weight ~ hindft, data = hares)

# tidy data frame of linear regression outputs
hindft_weight_lm_tidy <- tidy(hindft_weight_lm) 

# making it report-ready with equatiomatic
hindft_weight_lm_eq <- extract_eq(model = hindft_weight_lm)

# finding pearson's R for this relationship
pearsons_r <- cor.test(hares$hindft, hares$weight)
```

```{r, include = FALSE}
# exploring the linear regression results
summary(hindft_weight_lm)
plot(hindft_weight_lm)

```

\*\*\*\*2-3 sentences on what's up with the lm \#\# 5. Summary \#\# 5.
Summary \#\# 6. Citations

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017.
Snowshoe hare physical data in Bonanza Creek Experimental Forest:
1999-Present ver 22. Environmental Data Initiative.
<https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14>
(Accessed 2021-11-05).
